\bf Model Visualization Techniques for a Social Network Model

### 2017-08-22 paper.Rnw
done with last plot description (for now). take a break, then move on to the discussion.

<<alledgests, fig.height = 4, fig.cap="Visualizing all microsteps taken in 1,000 simulations from the model M1. The first wave network is shown at microstep 0, and the second wave of the network is shown as the last microstep for comparison. We see that it is rare for a microstep process to last longer than 150 steps, and also that the edges that appear past the 150th step tend to be in either the first wave or the second wave.">>=
# get a plot of all microsteps from w1 to w2 by edge
alledges <- expand.grid(from = paste0("V", 1:16), to = paste0("V", 1:16))
alledges$eid <- 1:nrow(alledges)

@@

sfmsall5 <- sfmsall4 %>% ungroup %>% add_row(ms = max(sfmsall4$ms) + 1, 
        eid = wave2$eid, 
        count = NA, 
        alpha = 1, 
        from = wave2$from,
        to = wave2$to,
        total = wave2$total, 
        plotorder = wave2$plotorder)

ggplot(data = sfmsall5) + 
  geom_tile(aes(x = ms, y = reorder(eid,-plotorder), fill = alpha)) +
  #scale_alpha_continuous("occurence %") + 
  scale_fill_gradient(low = 'grey90', high = '#0868ac', name = "occurrence %") + 
  labs(x = "Microstep No.", y = "Edges (ordered by total # occurrences)") + 
  ThemeNoNet + 
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(), panel.grid = element_blank())
@

\st{We also combine all 1,000 simulations into a visualization like the one shown in Figure~\ref{fig:allsteps}. We first assign each possible edge an edge ID number so that we can keep track of it throughout all the microsteps and all the simlations. Then, we count up the total number of times each edge appears in the microstep process in each of the 1,000 simulations for use as an ordering variable later. We also count up the number of times an edge occurs in each microstep number in the 1,000 simulations. Since the number of microsteps in the process varies, the number of times an edge occurs decreases as the microstep number increases. Next, we compute a proportion, which we call the occurrence percentage, which is the number of times the edge occurred in a microstep divided by 1,000. Finally, we visualize all this information together in Figure ~\ref{fig:alledgests}. In this plot, all possible edges are shown, and we see that every one of the $16 \times 15 = 240$ possible edges in the network occurs at some point in the simulation process. We also see, however, that the process struggles to focus in on the edges in the second wave of the data. Ideally, we would like to see more occurrences of the second wave of data generally. About half of the edges in wave two are in the bottom half when ordered by number of occurrences, meaning they aren't appearing as much as they would if the model was truly excellent at capturing the mechanisms of tie change in the network.}

>>>>

<<alledgests, fig.height = 4, fig.cap="Visualizing all microsteps taken in 1,000 simulations from the model M1. The occurrence percent is split up into groups to correspond with its distribution: only about 10\% of the edges appear more than 10\% of the time in the 1,000 simulations, while about 60\% appear less than 1\% of the time. The first wave network is shown at microstep 0, and the second wave of the network is shown as the last microstep for comparison. We see that it is rare for a microstep process to last longer than 150 steps, and also that the edges that appear past the 150th step tend to be in either the first wave or the second wave.">>=
# get a plot of all microsteps from w1 to w2 by edge
alledges <- expand.grid(from = paste0("V", 1:16), to = paste0("V", 1:16))
alledges$eid <- 1:nrow(alledges)

@@

sfmsall5 <- sfmsall4 %>% ungroup %>% add_row(ms = max(sfmsall4$ms) + 1, 
        eid = wave2$eid, 
        count = NA, 
        alpha = 1, 
        from = wave2$from,
        to = wave2$to,
        total = wave2$total, 
        plotorder = wave2$plotorder) 
sfmsall5$alphaFctr <- cut(sfmsall5$alpha, c(0,.01, .05, .10, .25, .5, .75, 1, 1.1))
ggplot(data = sfmsall5) + 
  geom_tile(aes(x = ms, y = reorder(eid,-plotorder), fill = alphaFctr)) +
  #scale_fill_gradient(low = 'grey90', high = '#0868ac', name = "occurrence %") + 
  scale_fill_brewer(palette = "Blues", name = "Occurrence %", labels = c("(0-1]%", "(1-5]%", "(5-10]%","(10-25]%", "(25-50]%", "(50-75]%", "(75-100]%")) + 
  labs(x = "Microstep No.", y = "Edges (ordered by total # occurrences)") + 
  ThemeNoNet + 
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(), panel.grid = element_blank())
@

\st{We also combine 1,000 simulations from model M1 into a visualization like the one shown in Figure~\ref{fig:allsteps}. We first assign each possible edge an edge ID number so that we can keep track of it throughout all the microsteps and all the simlations. Then, we count up the total number of times each edge appears in the microstep process in each of the 1,000 simulations for use as an ordering variable later. We also count up the number of times an edge occurs in each microstep number in the 1,000 simulations. Since the number of microsteps in the process varies, the number of times an edge occurs decreases as the microstep number increases. Next, we compute a proportion, which we call the occurrence percentage, which is the number of times the edge occurred in a microstep divided by 1,000. Finally, we visualize all this information together in Figure ~\ref{fig:alledgests}. In this plot, all possible edges are shown, and we see that every one of the $16 \times 15 = 240$ possible edges in the network occurs at some point in the simulation process. We also see, however, that the process struggles to focus in on the edges in the second wave of the data. Ideally, we would like to see more occurrences of the second wave of data generally. About half of the edges in wave two are in the bottom half when ordered by number of occurrences, meaning they aren't appearing as much as they would if the model was truly excellent at capturing the mechanisms of tie change in the network. In addition, there is not a distinct pattern or structure that we can see in this particular edge visualization. The edges that the model starts with appear more often in the first 50 or so microsteps than any other set of edges. This makes sense, but it would be better if the model explored the space more than it currently does. It could be that the simple model M1 isn't complicated enough for the friendship dynamics in the data, or it could be that the chains in the MCMCs "get stuck" when exploring the parameter space.}
